---
title: 'Final Project: Food and Nutrients Exploration'
author: "Group M: Xinyao Huang, Jiayi Song, Rou Xie, Juanyu Liu"
date: "2024-04-21"
output: ioslides_presentation
runtime: shiny
---


## Data Sources:

- **Supply Utilization Accounts Food and Diet 2010-2020**  
  Data from the Food and Agriculture Organization (FAO).  
  [FAO Stat Data](https://www.fao.org/faostat/en/#data/SUA)

- **National Health and Nutrition Examination Survey (NHANES) 2015-2016**  
  Data sourced from the Global Dietary Database.  
  [Global Dietary Database NHANES](https://www.globaldietarydatabase.org/management/microdata-surveys/718)

- **Food Consumption and Nutrient Intakes 2015-2018**  
  Data obtained from the USDA Economic Research Service.  
  [USDA ERS - Food Consumption and Nutrient Intakes](https://www.ers.usda.gov/data-products/food-consumption-and-nutrient-intakes/)


# Part 1. Explore US dataset

## Food Consumption and Nutrient Intakes
```{r, echo = FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(readr)
#P1
data <- read_csv("Total for categories 2015-2016 and 2017-2018.csv")

data$Food_group <- factor(data$Foodgroup, levels = unique(data$Foodgroup))

ggplot(data, aes(x = Food_group, y = Total, fill = Year)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  theme_minimal() +
  labs(title = 'Total for categories 2015-2016 and 2017-2018', x = 'Category', y = 'Total') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Sugar Intake Proportion
```{r, echo = FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(readr)
library(dplyr)
library(gridExtra)
library(scales)

data <- read_csv("sugar intake proportion.csv")

data <- data %>%
  filter(Group != "Total") %>%
  group_by(Year) %>%
  mutate(Proportion = Value / sum(Value)) %>%
  ungroup() %>%
  mutate(Label = ifelse(Proportion > 0.1, percent(Proportion), ""))

data_2015_2016 <- data %>% filter(Year == "2015-2016")
data_2017_2018 <- data %>% filter(Year == "2017-2018")

plot_pie <- function(data, title) {
  ggplot(data, aes(x = "", y = Proportion, fill = Group)) +
    geom_bar(stat = "identity", width = 1) +
    coord_polar("y", start = 0) +
    theme_void() +
    labs(title = title, fill = "Source") +
    geom_text(aes(label = Label), position = position_stack(vjust = 0.5), check_overlap = TRUE) +
    theme(
      legend.position = "bottom",  
      legend.direction = "vertical",  
      legend.key.width = unit(1, "cm"),  
      legend.text = element_text(size = 10)  
    )
}


p1 = plot_pie(data_2015_2016, "Sugar Intake Proportions 2015-2016")
p2 = plot_pie(data_2017_2018, "Sugar Intake Proportions 2017-2018")

grid.arrange(p1, p2, nrow = 1)



```


## Comparison by groups
```{r, echo = FALSE,message=FALSE, warning=FALSE}

library(ggplot2)
library(readr)

data <- read_csv("people_group_sugarintake.csv")

data$Group <- factor(data$Group, levels = unique(data$Group))

ggplot(data, aes(x = Group, y = Total, fill = Year)) +
  geom_bar(stat = 'identity', position = position_dodge()) +
  scale_fill_manual(values = c("2015-2016" = "blue", "2017-2018" = "orange")) +
  theme_minimal() +
  labs(x = "Group", y = "Total Sugar Intake (tsp)", title = "Total Sugar Intake by Group and Year") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


## Home vs. Away 
```{r, echo = FALSE, message=FALSE, warning=FALSE}
data <- read_csv("peope_group_location_sugar.csv")

data$Group <- factor(data$Group, levels = unique(data$Group))
data$Location <- factor(data$Location, levels = c("At home", "Away from home"))

# ggplot(data, aes(x = Group, y = Total, fill = Location)) +
#   geom_bar(stat = 'identity', position = position_dodge()) +
#   scale_fill_manual(values = c("At home" = "blue", "Away from home" = "orange")) +
#   theme_minimal() +
#   labs(title = 'Comparative Sugar Intake by Group at Home vs. Away Across (2015-2016,2017-2018)',
#        x = 'Group',
#        y = 'Total Sugar Intake (tsp)',
#        fill = 'Location') +
#   theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
#         plot.title = element_text(size = rel(1), hjust = 0.5),
#         legend.position = "bottom") +
#   facet_wrap(~Year, ncol = 1)

data$Group <- factor(data$Group, levels = unique(data$Group))


ggplot(data, aes(x = Group, y = Total, fill = Location)) +
  geom_bar(stat = 'identity', position = position_dodge(width = 0.9), alpha = 0.7) +
  scale_fill_manual(values = c("At home" = "blue", "Away from home" = "orange")) +
  theme_minimal() +
  labs(title = 'Comparative Sugar Intake by Group at Home vs. Away (2015-2016,2017-2018)',
       x = 'Group',
       y = 'Total Sugar Intake (tsp)',
       fill = 'Location') +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
        plot.title = element_text(size = rel(1), hjust = 0.5),
        legend.position = "bottom") +
  facet_grid(. ~ Year)


```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse) 
library(dplyr)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
diet_data_part1 <- read.csv("DietData_part1.csv")
diet_data_part2 <- read.csv("DietData_part2.csv")
participant_data <- read.csv("ParticipantData.csv")
diet_data <- bind_rows(diet_data_part1, diet_data_part2)
combined_data <- inner_join(diet_data, participant_data, by = "id")

#Clean data
combined_data <- combined_data %>%
  filter(!is.na(totalsugars), !is.na(pa), !is.na(age.x), !is.na(sex.x))

#Add labels
race_labels <- c("Mexican American", "Other Hispanic", "White", "Black", "Asian", "Other Race")
sex_labels <- c("Male", "Female")
combined_data$race_label <- race_labels[combined_data$race]
combined_data$sex_label <- sex_labels[combined_data$sex.x]
```

## Sugar & Race
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Total Sugar Intake divide by race
avg_sugar_by_race <- combined_data %>%
  group_by(race) %>%
  summarize(avg_sugar = mean(totalsugars, na.rm = TRUE))

# Line Chart
ggplot(avg_sugar_by_race, aes(x = race, y = avg_sugar, fill = race_labels)) +
  geom_bar(stat = "identity") +
  labs(title = "Average Sugar Intake by Race", x = "Race", y = "Average Sugar Intake (g)") +
  scale_fill_discrete(name = "Race", labels = race_labels) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Scatter plot
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# A scatter plot shows the relationship between sugar intake and physical activity
ggplot(combined_data, aes(x = totalsugars, y = pa)) +
  geom_point(aes(color = race_label), alpha = 0.6) +
  geom_smooth(method = "lm", color = "blue") +
  labs(title = "Sugar Intake vs Physical Activity", x = "Total Sugar Intake (g)", y = "Physical Activity (MET-min/week)") +
  theme_minimal()
```


## Analyze Total sugar intake and Sex
```{r, echo=FALSE, message=FALSE, warning=FALSE}
avg_sugar_by_sex <- combined_data %>%
  group_by(sex.x) %>%
  summarize(avg_sugar = mean(totalsugars, na.rm = TRUE))

# ggplot(avg_sugar_by_sex, aes(x = sex.x, y = avg_sugar, fill = sex_labels)) +
#   geom_bar(stat = "identity") +
#   labs(title = "Average Sugar Intake by Male and female", x = "Sex", y = "Average Sugar Intake (g)") +
#   scale_fill_discrete(name = "Sex", labels = sex_labels) +
#   theme_minimal() +
#   theme(legend.position = "bottom")

ggplot(combined_data, aes(x = totalsugars, y = pa)) +
  geom_point(aes(color = sex_label), alpha = 0.6) +
  geom_smooth(method = "lm", color = "blue") +
  labs(title = "Sugar Intake vs Physical Activity", x = "Total Sugar Intake (g)", y = "Physical Activity (MET-min/week)") +
  theme_minimal()
```


# Part 2. Explore global dataset: Supply Utilization Accounts Food and Diet 2010-2020

```{r, echo=FALSE, message=FALSE, warning=FALSE}
df <- read.csv("Supply_Utilization_Accounts_Food_and_Diet_E_All_Data_(Normalized).csv", fileEncoding = "ISO-8859-1")
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(shiny)
library(DT)
library(dplyr)
library(shinyWidgets)
library(ggplot2)
library(plotly)
```
```{r, echo = FALSE}
df <- df[, c("Area", "Food.Group","Indicator", "Year", "Unit", "Value")]
```

## Multiple Search by Countries
```{r, echo=FALSE, message=FALSE, warning=FALSE}
nutrients <- unique(df$Indicator)
foodgroups <- unique(df$Food.Group)

ui <- fluidPage(
  titlePanel("Average Daily Nutrient Availability by Country"),
  sidebarLayout(
    sidebarPanel(
      selectizeInput("nutrient", "Select Nutrient", choices = nutrients, multiple = FALSE),
      selectizeInput("foodgroup", "Select Food Group", choices = foodgroups, multiple = FALSE),
      sliderInput("year_range", "Select Year Range", min = 2010, max = 2020, value = c(2010,2020),step = 1),
      textAreaInput("countries", "Enter countries (one per line):", value = "", width = "400px", height = "200px"),
      actionButton("submit", "Submit")
    ),
    mainPanel(
      DTOutput("filtered_table"),
      plotlyOutput("line_plot")
    )
  )
)
years <- c(2010, 2012, 2014, 2016, 2018, 2020)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
server <- function(input, output) {
  
  filtered_data <- reactive({
    countries <- unlist(strsplit(input$countries, "\n"))
    filtered <- df[df$Area %in% countries, ]
    
    if (!is.null(input$year_range)) {
      filtered <- filtered %>%
        filter(Year >= input$year_range[1] & Year <= input$year_range[2])
    }
    
    if (!is.null(input$nutrient)) {
      filtered <- filtered %>%
        filter(Indicator %in% input$nutrient)
    }
    
    if (!is.null(input$foodgroup)) {
      filtered <- filtered %>%
        filter(Food.Group %in% input$foodgroup)
    }
    
    return(filtered)
  })
  
  output$filtered_table <- renderDT({
    datatable(filtered_data(), options = list(pageLength = 10))
  })
  
  output$line_plot <- renderPlotly({
    plot <- ggplot(filtered_data(), aes(x = Year, y = Value, color = Area)) +
      geom_line() +
      labs(x = "time", y = "value", title = "Average Daily Nutrient Availability Trend") +
      theme_minimal()+
      scale_x_continuous(breaks = years)
    ggplotly(plot)
  })
  
}

shinyApp(ui = ui, server = server)
```


