---
title: 'Project: Food and Nutrients Exploration'
author: "Group M"
date: "2024-04-21"
output: html_document
---


# Part 1. Explore US dataset: 


```{r}

library(ggplot2)
library(readr)
#P1
data <- read_csv("Total for categories 2015-2016 and 2017-2018.csv")

data$Food_group <- factor(data$Foodgroup, levels = unique(data$Foodgroup))

ggplot(data, aes(x = Food_group, y = Total, fill = Year)) +
  geom_bar(stat = 'identity', position = 'dodge') +
  theme_minimal() +
  labs(title = 'Total for categories 2015-2016 and 2017-2018', x = 'Category', y = 'Total') +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```


This Graph shows the total quantities for various food categories over two different time periods (2015-2016 and 2017-2018), allowing for a direct comparison. The data can help examine the relationship between eating habits and health. If some less healthy food categories show an upward trend, it could indicate a potential risk to public health.


```{r}

library(ggplot2)
library(readr)
library(dplyr)
library(gridExtra)
library(scales)

data <- read_csv("sugar intake proportion.csv")

data <- data %>%
  filter(Group != "Total") %>%
  group_by(Year) %>%
  mutate(Proportion = Value / sum(Value)) %>%
  ungroup() %>%
  mutate(Label = ifelse(Proportion > 0.1, percent(Proportion), ""))

data_2015_2016 <- data %>% filter(Year == "2015-2016")
data_2017_2018 <- data %>% filter(Year == "2017-2018")

plot_pie <- function(data, title) {
  ggplot(data, aes(x = "", y = Proportion, fill = Group)) +
    geom_bar(stat = "identity", width = 1) +
    coord_polar("y", start = 0) +
    theme_void() +
    labs(title = title, fill = "Source") +
    geom_text(aes(label = Label), position = position_stack(vjust = 0.5), check_overlap = TRUE) +
    theme(
      legend.position = "bottom",  
      legend.direction = "vertical",  
      legend.key.width = unit(1, "cm"),  
      legend.text = element_text(size = 10)  
    )
}


p1 = plot_pie(data_2015_2016, "Sugar Intake Proportions 2015-2016")
p2 = plot_pie(data_2017_2018, "Sugar Intake Proportions 2017-2018")

grid.arrange(p1, p2, nrow = 1)



```

The two pie charts illustrate the proportions of sugar intake from different sources and compare the 2015-2016 and 2017-2018 periods.
The "At home" segment is the largest, accounting for approximately 69.86%, indicating that most sugar intake occurs in home-based settings. Other sources (such as fast food, restaurants, etc.) hold smaller shares. 2017-2018 Pie Chart:
The "At home" segment remains the largest, accounting for approximately 71.61%, indicating an increase in sugar intake from home-based sources. Other sources show slight changes but maintain relatively smaller proportions.


```{r}

library(ggplot2)
library(readr)

data <- read_csv("people_group_sugarintake.csv")

data$Group <- factor(data$Group, levels = unique(data$Group))

ggplot(data, aes(x = Group, y = Total, fill = Year)) +
  geom_bar(stat = 'identity', position = position_dodge()) +
  scale_fill_manual(values = c("2015-2016" = "blue", "2017-2018" = "orange")) +
  theme_minimal() +
  labs(x = "Group", y = "Total Sugar Intake (tsp)", title = "Total Sugar Intake by Group and Year") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

The bar graph comparing total sugar intake among different groups for two periods: 2015-2016 and 2017-2018. 
It visualizes how sugar intake varies by group and year, with bars grouped by category and distinguished by color.
In general, the bars for 2017-2018 (orange) are higher than those for 2015-2016 (blue), suggesting an increase in total sugar intake across most groups.


```{r}
data <- read_csv("peope_group_location_sugar.csv")

data$Group <- factor(data$Group, levels = unique(data$Group))
data$Location <- factor(data$Location, levels = c("At home", "Away from home"))

ggplot(data, aes(x = Group, y = Total, fill = Location)) +
  geom_bar(stat = 'identity', position = position_dodge()) +
  scale_fill_manual(values = c("At home" = "blue", "Away from home" = "orange")) +
  theme_minimal() +
  labs(title = 'Comparative Sugar Intake by Group at Home vs. Away Across (2015-2016,2017-2018)',
       x = 'Group',
       y = 'Total Sugar Intake (tsp)',
       fill = 'Location') +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
        plot.title = element_text(size = rel(1), hjust = 0.5),
        legend.position = "bottom") +
  facet_wrap(~Year, ncol = 1)

data$Group <- factor(data$Group, levels = unique(data$Group))


ggplot(data, aes(x = Group, y = Total, fill = Location)) +
  geom_bar(stat = 'identity', position = position_dodge(width = 0.9), alpha = 0.7) +
  scale_fill_manual(values = c("At home" = "blue", "Away from home" = "orange")) +
  theme_minimal() +
  labs(title = 'Comparative Sugar Intake by Group at Home vs. Away (2015-2016,2017-2018)',
       x = 'Group',
       y = 'Total Sugar Intake (tsp)',
       fill = 'Location') +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1),
        plot.title = element_text(size = rel(1), hjust = 0.5),
        legend.position = "bottom") +
  facet_grid(. ~ Year)


```


The chart illustrates a comparison of sugar intake at home and away from home for different groups during two periods: 2015-2016 and 2017-2018. 
For most groups, sugar intake at home is significantly higher than away from home, especially for adults and middle-income households. This suggests that home eating habits play an essential role in sugar intake.
The bar chart allows you to compare different groups in terms of sugar intake at home and away from home.
In both 2015-2016 and 2017-2018, "At home" generally has higher sugar intake than "Away from home," suggesting that home-based consumption contributes more to total sugar intake.
Among the groups, "adult" and "middle-income households" have the highest total sugar intake, while "low-income households" and "elder" have lower values.
The trend shows that from 2015-2016 to 2017-2018, sugar intake at home has generally increased, indicating a shift in dietary patterns or consumption behaviors.



# Part 2. Explore global dataset: Supply Utilization Accounts Food and Diet 2010-2020
```{r}
df <- read.csv("Supply_Utilization_Accounts_Food_and_Diet_E_All_Data_(Normalized).csv")
```

```{r}
library(shiny)
library(DT)
library(dplyr)
library(shinyWidgets)
library(ggplot2)
library(plotly)
```
```{r}
df <- df[, c("Area", "Food.Group","Indicator", "Year", "Unit", "Value")]
```

## Multiple Search by Countries
```{r}
nutrients <- unique(df$Indicator)
foodgroups <- unique(df$Food.Group)

ui <- fluidPage(
  titlePanel("Average Daily Nutrient Availability by Country"),
  sidebarLayout(
    sidebarPanel(
      selectizeInput("nutrient", "Select Nutrient", choices = nutrients, multiple = FALSE),
      selectizeInput("foodgroup", "Select Food Group", choices = foodgroups, multiple = FALSE),
      sliderInput("year_range", "Select Year Range", min = 2010, max = 2020, value = c(2010,2020),step = 1),
      textAreaInput("countries", "Enter countries (one per line):", value = "", width = "400px", height = "200px"),
      actionButton("submit", "Submit")
    ),
    mainPanel(
      DTOutput("filtered_table"),
      plotlyOutput("line_plot")
    )
  )
)

```

```{r}
years <- c(2010, 2012, 2014, 2016, 2018, 2020)
```

```{r}
server <- function(input, output) {
  
  filtered_data <- reactive({
    countries <- unlist(strsplit(input$countries, "\n"))
    filtered <- df[df$Area %in% countries, ]
    
    if (!is.null(input$year_range)) {
      filtered <- filtered %>%
        filter(Year >= input$year_range[1] & Year <= input$year_range[2])
    }
    
    if (!is.null(input$nutrient)) {
      filtered <- filtered %>%
        filter(Indicator %in% input$nutrient)
    }
    
    if (!is.null(input$foodgroup)) {
      filtered <- filtered %>%
        filter(Food.Group %in% input$foodgroup)
    }
    
    return(filtered)
  })
  
  output$filtered_table <- renderDT({
    datatable(filtered_data(), options = list(pageLength = 10))
  })
  
  output$line_plot <- renderPlotly({
    plot <- ggplot(filtered_data(), aes(x = Year, y = Value, color = Area)) +
      geom_line() +
      labs(x = "time", y = "value", title = "Average Daily Nutrient Availability Trend") +
      theme_minimal()+
      scale_x_continuous(breaks = years)
    ggplotly(plot)
  })
  
}

shinyApp(ui = ui, server = server)
```


